---
title: "US Honey Production"
author: "Tristan Hamilton"
date: "2024-03-05"
categories: [Blog Post 3]
---


```{r include=FALSE} 
library(tidyverse)
library(here)
library(knitr)
us_honey <- read_csv(here("data/honeyproduction.csv"))
```


Ideas:
-Examine trends in the total production of honey over the years for each state

-Compare the average yield per colony across different states / Explore variations in honey production among states

-

Variables: 

state = US State

numcol = Number of colonies producing honey: This variable indicates the total number of bee colonies involved in honey production within a specific area or timeframe. It's an essential metric for understanding the scale of honey production operations.

yieldpercol = Yield per colony: This variable represents the average amount of honey produced by each colony. It helps assess the efficiency and productivity of beekeeping practices, as well as the health of bee populations.

totalprod = Total production of honey: This variable shows the overall quantity of honey produced within the dataset's scope. It's a fundamental measure of honey output and is influenced by factors such as colony count and yield per colony.

stocks = Honey stocks: This variable refers to the amount of honey stored or available for use at a given point in time. It's crucial for understanding supply dynamics within the honey market and can impact pricing and availability.

priceperlb = Average price per pound: This variable indicates the average cost of honey per unit weight (typically per pound) within the dataset's context. It reflects market demand, supply conditions, and factors affecting pricing, such as seasonality and regional variations.

prodvalue = Average production value: This variable represents the average monetary worth of honey production, considering both quantity and price. It provides insights into the economic significance of honey production and its contribution to the overall agricultural sector.

year = 1998 - 2012

### Introduction:

<https://github.com/rfordatascience/tidytuesday/blob/master/data/2018/2018-05-21/week8_honey_production.zip>.
```{r}
kable(head(us_honey))
```


### Part 1:

```{r}
us_honey |> group_by(state) |>
  summarise(totalhoneyprod = sum(totalprod)) |>
  arrange(desc(totalhoneyprod)) |>
  slice(1:10)
```

```{r}
## Take the top 6 states from the prior table
top4_df <- us_honey |> 
  filter(state == "ND" | state == "CA" | state == "SD" | state == "FL")
```

First look at North Dakota:
```{r}
ggplot(nd_df, aes(x = year, y = totalprod, colour = state)) +
  geom_line(colour = "turquoise3", linewidth = 1.3) +
  geom_area(fill = "turquoise3", alpha = 0.3) +
  labs(x = "Year (1998-2012)",
       y = "North Dakota Total Honey Production") +
  theme_minimal()
```

Then with the rest of the top 6 states:
```{r}
ggplot(top4_df, aes(x = year, y = totalprod, colour = state)) +
  geom_line(linewidth = 1.4) +
  labs(x = "Year",
       y = "Total Honey Production") +
  theme_minimal()
```
-Here we can look at the total honey production among the top states over the time period 1998 to 2012. We can see that ND has stood out as a industry leader. 

```{r}
honey_diff <- us_honey |> filter(year == "1998" | year == "2012") |>
  group_by(state) |>
  summarise(diff_prod = diff(totalprod))

honey_diff <- honey_diff |> mutate(change = if_else(diff_prod > 0,
                                                    true = "increase",
                                                    false = "decrease"))
honey_joined <- left_join(us_honey, honey_diff)

ggplot(honey_joined, aes(x = year, y = totalprod)) +
  geom_line(aes(group = state, colour = change), alpha = 0.6) +
  theme_minimal() +
  scale_colour_viridis_d()
```


Price movements over the time period:
```{r}
ggplot(top4_df, aes(x = year, y = priceperlb, colour = state)) +
  geom_line(linewidth = 0.9) +
  labs(x = "Year",
       y = "Price per Pound") +
  theme_minimal()
```
Shows that price movements in honey are consistent across US states, all follow the same trend 

### Part 2:

On the surface, some of the variables seem very similar, such as total production, number of colonies, stocks, etc., it would be interesting to look at correlation matrix between them. 
```{r}
corr_honey <- us_honey |>
  select(totalprod, numcol, yieldpercol, stocks)

corr_matrix <- cor(corr_honey)

# install.packages("corrplot")
library(corrplot)

corrplot(corr_matrix, method = "color", addCoef.col = "green")
```
Total production of honey correlates highly with the number of colonies and honey stocks for each state. That is, higher production is associated with a larger number of colonies and larger honey stocks. 

we can also use ggpairs to look at the correlation between each pair of variables. It shows the pairwise relationships with scatterplots. 
```{r}
library(GGally)
ggpairs(data = us_honey, columns = 2:5)
```


### Conclusion:

### Connection to Class:
